Meta RL analyses

### 2019-04-30 meta_RL_analyses.Rmd
Baselining for meta + more aesthetics

@@

```{r}
name_var = function(var) { # var names for plotting
  case_when(var %in% c("is_new", "is_new_game") ~ "Held out?",
            var == "game_type" ~ "Game",
            var %in% c("run_type", "named_run_type") ~ "Run type",
#            var == "reward" ~ "Reward",
            grepl("reward", var) ~ "Average Reward",
            T ~ var)
  }

summary_plot = function(data,  x_var, y_var, color_var, optimal_baseline_data=NULL, other_baseline_data=NULL, palette="Set2") {
  p = ggplot(data, aes_string(x=x_var, y=y_var, color=color_var)) +  
    geom_hline(yintercept=0., linetype=2, alpha=0.5) +
    geom_point(stat="summary", fun.y=mean, size=2) +
    scale_color_manual(values=c("#6560a3", "#f37f22"))

@@

  if (!is.null(other_baseline_data)) {
    p = p + 
      geom_spoke(data=other_baseline_data,
                 aes(x=game_type, y=expected_reward, angle=0, radius=1),
                 linetype=2, alpha=0.5,
                 position=position_nudge(x=-0.5)) 
  }

  p = p + 
    geom_errorbar(stat="summary", 
                  fun.data="mean_cl_boot",
                  width=0.25) +
#    scale_color_brewer(palette=palette) +
    labs(x=name_var(x_var), y=name_var(y_var)) +
    guides(color=guide_legend(title=NULL))

@@

```{r}
baseline_data = read_csv("baseline_data.csv") %>%
  mutate(game_type = str_extract(game, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game),
         black_valuable = grepl("bv_1", game),
         suits_rule = grepl("sr_1", game))
```
```{r}
optimal_baselines = baseline_data %>% 
  filter(policy == "optimal") %>%
  group_by(game_type) %>%
  summarize(expected_reward = mean(expected_reward)) %>%
  ungroup()
```

@@

```{r}
most_correlated_baselines = crossing(run_type=result_subdirs, baseline_data) %>%
  mutate(is_new_game = check_if_new_game(run_type, game)) %>%
  filter((!is_new_game & policy == "optimal") | (policy != "optimal" & !check_if_new_game(run_type, policy))) %>%
  group_by(run_type, game, is_new_game, game_type, losers, black_valuable, suits_rule) %>%
  summarize(optimal_expected_reward = max(expected_reward)) %>%
  group_by(run_type, game_type, is_new_game) %>%
  summarize(expected_reward = mean(optimal_expected_reward)) %>%
  mutate(is_new_game = factor(is_new_game, levels=c(F, T), labels=c("Trained", "Held out")),
         named_run_type = name_run_type(run_type))

@@

reward_d = reward_d %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one | is_only_one_ho)) %>%
  gather(game_str, reward, starts_with("game")) %>%
  mutate(game_type = str_extract(game_str, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game_str),
         black_valuable = grepl("bv_1", game_str),
         suits_rule = grepl("sr_1", game_str),
         is_new_game = check_if_new_game(run_type, game_str),
         is_new_game = factor(is_new_game, labels=c("Trained", "Held out")))
```

```{r}
avg_reward_d = reward_d %>%
  select(-game_str) %>%
  filter(epoch == 0) %>%
  group_by(run_type, game_type, is_new_game) %>%
  summarize(avg_reward = mean(reward), sd_reward=sd(reward),
            se_reward = sd_reward/sqrt(n()), 
            ci_lo = avg_reward - 1.96*se_reward, 
            ci_hi = avg_reward + 1.96*se_reward)
```

@@

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0,
                      run_type == "basic"),
             "game_type", "reward", "is_new_game",
             optimal_baselines,
             most_correlated_baselines %>%
               filter(run_type == "basic",
                      is_new_game == "Held out")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0,
                      run_type == "basic_random_holdout_longer"),
             "game_type", "reward", "is_new_game",
             optimal_baselines,
             most_correlated_baselines %>%
               filter(run_type == "basic_random_holdout_longer",
                      is_new_game == "Held out")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0),
             "game_type", "reward", "is_new_game",
             optimal_baselines,
             most_correlated_baselines %>%
               filter(is_new_game == "Held out")) +
  facet_wrap(. ~ named_run_type) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ggsave("writing/figures/basic_meta_learning.png", width=5, height=3.5)
```

@@

@@

meta_d = meta_d %>%
  filter(epoch == 0) %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one)) %>%
  gather(map_str, reward, -c(run, run_type, named_run_type, epoch), -starts_with("is_")) %>%
  separate(map_str, c("meta_task", "source", "target"), sep="[\\.]+") %>%
  mutate(is_new = check_if_new_game(run_type, source) | check_if_new_game(run_type, target),
         is_new = factor(is_new, labels=c("Trained", "Held out")))
```

```{r}
avg_meta_d = meta_d %>%
  group_by(run_type, named_run_type, run, meta_task, is_new) %>%
  summarize(avg_reward = mean(reward)) %>%
  group_by(run_type, named_run_type, meta_task, is_new) %>%
  summarize(avg_avg_reward = mean(avg_reward), sd_avg_reward = sd(avg_reward),
            se_reward = sd_avg_reward/sqrt(n()), 
            ci_lo = avg_avg_reward - 1.96*se_reward, 
            ci_hi = avg_avg_reward + 1.96*se_reward)

@@

```{r}
summary_plot(meta_d %>%
               filter(epoch == 0,
                      run_type %in% c("basic", "basic_random_holdout_longer")),
             "meta_task", "reward", "is_new") +
  facet_wrap(. ~ named_run_type)  +
  guides(color=guide_legend(title=NULL)) +
  labs(x="Game type", y="Average reward") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

>>>>

name_game_type = function(game) {
  case_when(game == "high_card" ~ "High card",
            game == "straight_flush" ~ "Straight flush",
            game == "match" ~ "Match",
            game == "sum_under" ~ "Blackjack",
            game == "pairs_and_high" ~ "Pairs",
            T ~ game)
}

name_meta_task = function(meta_task) {
  case_when(meta_task == "toggle_losers" ~ "Losers",
            meta_task == "toggle_suits_rule" ~ "Suits > rank",
            meta_task == "toggle_black_valuable" ~ "Switch suits",
            T ~ meta_task)
}

@@

```{r}
name_var = function(var) { # var names for plotting
  case_when(var %in% c("is_new", "is_new_game") ~ "Held out?",
            var %in% c("game_type", "named_game_type") ~ "Game",
            var %in% c("run_type", "named_run_type") ~ "Run type",
#            var == "reward" ~ "Reward",
            grepl("reward", var) ~ "Average Reward",
            T ~ var)
  }

summary_plot = function(data,  x_var, y_var, color_var, optimal_baseline_data=NULL, other_baseline_data=NULL, palette="Set2") {

  p = ggplot(data, aes_string(x=x_var, y=y_var, color=color_var)) +  
    geom_hline(yintercept=0., linetype=2, alpha=0.5) +
    geom_point(stat="summary", fun.y=mean, size=2) +
    scale_color_manual(values=c("#555093", "#f38f22"))

@@

  if (!is.null(other_baseline_data)) {
    # get rid of runs that aren't in data
    run_types_included = data %>% pull(named_run_type) %>% unique()
    other_baseline_data = other_baseline_data %>%
      filter(named_run_type %in% run_types_included)

    p = p + 
      geom_spoke(data=other_baseline_data,
                 aes_string(x=x_var, y="expected_reward", angle=0, radius=1),
                 linetype=2, alpha=0.5,
                 position=position_nudge(x=-0.5)) 
  }

  p = p + 
    geom_point(stat="summary", fun.y=mean, size=2) +
    geom_errorbar(stat="summary", 
                  fun.data="mean_cl_boot",
                  width=0.4) +
#    scale_color_brewer(palette=palette) +
    labs(x=name_var(x_var), y=name_var(y_var)) +
    guides(color=guide_legend(title=NULL))

@@

```{r}
baseline_data = read_csv("baseline_data.csv") %>%
  mutate(game_type = str_extract(game, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game),
         black_valuable = grepl("bv_1", game),
         suits_rule = grepl("sr_1", game),
         named_game_type = name_game_type(game_type))
```
```{r}
optimal_baselines = baseline_data %>% 
  filter(policy == "optimal") %>%
  group_by(game_type, named_game_type) %>%
  summarize(expected_reward = mean(expected_reward)) %>%
  ungroup() %>%
  arrange(named_game_type)
```

@@

```{r}
most_correlated_baselines = crossing(run_type=result_subdirs, baseline_data) %>%
  mutate(is_new_game = check_if_new_game(run_type, game)) %>%
  filter((!is_new_game & policy == "optimal") | (policy != "optimal" & !check_if_new_game(run_type, policy))) %>%
  group_by(run_type, game, is_new_game, game_type, losers, black_valuable, suits_rule) %>%
  summarize(optimal_expected_reward = max(expected_reward)) %>%
  group_by(run_type, game_type, is_new_game) %>%
  summarize(expected_reward = mean(optimal_expected_reward)) %>%
  mutate(is_new_game = factor(is_new_game, levels=c(F, T), labels=c("Trained", "Held out")),
         named_run_type = name_run_type(run_type),
         named_game_type = name_game_type(game_type))

@@

reward_d = reward_d %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one | is_only_one_ho)) %>%
  gather(game_str, reward, starts_with("game")) %>%
  mutate(game_type = str_extract(game_str, "high_card|match|straight_flush|pairs_and_high|sum_under"),
         losers = grepl("l_1", game_str),
         black_valuable = grepl("bv_1", game_str),
         suits_rule = grepl("sr_1", game_str),
         is_new_game = check_if_new_game(run_type, game_str),
         is_new_game = factor(is_new_game, labels=c("Trained", "Held out")),
         named_game_type = name_game_type(game_type))
```

```{r}
avg_reward_d = reward_d %>%
  select(-game_str) %>%
  filter(epoch == 0) %>%
  group_by(run_type, game_type, named_game_type, is_new_game) %>%
  summarize(avg_reward = mean(reward), sd_reward=sd(reward),
            se_reward = sd_reward/sqrt(n()), 
            ci_lo = avg_reward - 1.96*se_reward, 
            ci_hi = avg_reward + 1.96*se_reward)
```

@@

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0,
                      run_type == "basic"),
             "named_game_type", "reward", "is_new_game",
             optimal_baselines,
             most_correlated_baselines %>%
               filter(run_type == "basic",
                      is_new_game == "Held out")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0,
                      run_type == "basic_random_holdout_longer"),
             "named_game_type", "reward", "is_new_game",
             optimal_baselines,
             most_correlated_baselines %>%
               filter(run_type == "basic_random_holdout_longer",
                      is_new_game == "Held out")) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}
summary_plot(reward_d %>%
               filter(epoch == 0,
                      run_type %in% c("basic", "basic_random_holdout_longer")),
             "named_game_type", "reward", "is_new_game",
             optimal_baselines,
             most_correlated_baselines %>%
               filter(is_new_game == "Held out")) +
  facet_wrap(. ~ named_run_type) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("writing/figures/basic_meta_learning.png", width=5, height=3)
```

@@

```{r}
meta_tasks = c("toggle_losers", "toggle_suits_rule", "toggle_black_valuable")
meta_baseline_unmapped_d = crossing(meta_task=meta_tasks,
                                    run_type=result_subdirs,
                                    baseline_data) %>%
  filter(case_when(meta_task == "toggle_losers" ~ game == gsub("l_0", "l_1", policy) | game == gsub("l_1", "l_0", policy), 
                   meta_task == "toggle_black_valuable" ~ game == gsub("bv_0", "bv_1", policy) | game == gsub("bv_1", "bv_0", policy), 
                   meta_task == "toggle_suits_rule" ~ game == gsub("sr_0", "sr_1", policy) | game == gsub("sr_1", "sr_0", policy), 
                   T ~ F)) %>%
  mutate(named_run_type = name_run_type(run_type),
         named_meta_task = name_meta_task(meta_task),
         is_new = check_if_new_game(run_type, game) | check_if_new_game(run_type, policy),
         is_new = factor(is_new, labels=c("Trained", "Held out"))) %>%
  group_by(run_type, named_run_type, meta_task, named_meta_task, is_new)  %>%
  summarize(expected_reward = mean(expected_reward))

@@

meta_d = meta_d %>%
  filter(epoch == 0) %>%
  mutate(is_su_ho=grepl("su_heldout", run_type),
         is_only_one=grepl("only_one$", run_type),
         is_only_one_ho=grepl("only_one_ho", run_type),
         is_random_ho=grepl("random", run_type),
         is_default=!(is_su_ho | is_random_ho | is_only_one)) %>%
  gather(map_str, reward, -c(run, run_type, named_run_type, epoch), -starts_with("is_")) %>%
  separate(map_str, c("meta_task", "source", "target"), sep="[\\.]+") %>%
  mutate(is_new = check_if_new_game(run_type, source) | check_if_new_game(run_type, target),
         is_new = factor(is_new, labels=c("Trained", "Held out")),
         named_meta_task = name_meta_task(meta_task))
```

```{r}
avg_meta_d = meta_d %>%
  group_by(run_type, named_run_type, run, meta_task, named_meta_task, is_new) %>%
  summarize(avg_reward = mean(reward)) %>%
  group_by(run_type, named_run_type, meta_task, named_meta_task, is_new) %>%
  summarize(avg_avg_reward = mean(avg_reward), sd_avg_reward = sd(avg_reward),
            se_reward = sd_avg_reward/sqrt(n()), 
            ci_lo = avg_avg_reward - 1.96*se_reward, 
            ci_hi = avg_avg_reward + 1.96*se_reward)

@@

```{r}
summary_plot(meta_d %>%
               filter(epoch == 0,
                      meta_task %in% c("toggle_losers", "toggle_black_valuable"),
                      run_type %in% c("basic", "basic_random_holdout_longer")),
             "named_meta_task", "reward", "is_new",
             other_baseline_data = meta_baseline_unmapped_d %>%
               filter(is_new == "Held out",
                      meta_task %in% c("toggle_losers", "toggle_black_valuable"))) +
  facet_wrap(. ~ named_run_type)  +
  guides(color=guide_legend(title=NULL)) +
  labs(x="Game type", y="Average reward") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
